---
title: "Cyclistic Case Study"
author: "Brayden Balla"
date: "2023-08-06"
output:
  pdf_document: default
  html_document: default
---


##### Loading packages I frequently use

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)

```

##### Importing my data 

```{r Loading Data}

June_2022 <- read_csv("202206-divvy-tripdata.csv")
July_2022 <- read_csv("202207-divvy-tripdata.csv")
August_2022 <- read_csv("202208-divvy-tripdata.csv")
September_2022 <- read_csv("202208-divvy-tripdata.csv") 
October_2022 <- read_csv("202210-divvy-tripdata.csv") 
November_2022 <- read_csv("202211-divvy-tripdata.csv")
December_2022 <- read_csv("202212-divvy-tripdata.csv")

January_2023 <- read_csv("202301-divvy-tripdata.csv")
Febuary_2023 <- read_csv("202302-divvy-tripdata.csv")
March_2023 <- read_csv("202303-divvy-tripdata.csv")
April_2023 <- read_csv("202304-divvy-tripdata.csv")
May_2023 <- read_csv("202305-divvy-tripdata.csv")
June_2023 <- read_csv("202306-divvy-tripdata.csv")

```


##### April_2023, June_2023, June_2022, August_2022, September_2022 have missing values for columns:
##### "start_station_name, start_station_id, end_station_name, end_station_id."

```{r Structure Preview}

str(April_2023)
str(March_2023)

```

##### It appears as though the data is formatted correctly. 

```{r Final structure preview}

colnames(April_2023)
colnames(March_2023)
str(Febuary_2023$rideable_type)

```

##### I want to know how many distinct values I have for rideable types and members/casual.
##### Rideable Types are electric bike, classic bike, docked bike. There are options for members and casual riders.  

```{r distinct values}

unique(April_2023$rideable_type)
unique(April_2023$member_casual)
unique(May_2023$rideable_type)
unique(May_2023$member_casual)  

```

##### Changing the original tables into tables for the final 2 quarters of 2022 and the first 2 of 2023.
##### There are separate tables for members and casual riders of each table.

```{r}
April2023_Casual <- April_2023 %>%
  filter(member_casual == 'casual')
April2023_Member <- April_2023 %>%
  filter(member_casual == 'member')
May2023_Casual <- May_2023 %>%
  filter(member_casual == 'casual')     ##Q2 2023
May2023_Member <- May_2023 %>%
  filter(member_casual == 'member')
June2023_Casual <- June_2023 %>%
  filter(member_casual == 'casual')
June2023_Member <- June_2023 %>%
  filter(member_casual == 'member')

January2023_Casual <- January_2023 %>%
  filter(member_casual == 'casual')
January2023_Member <- January_2023 %>%
  filter(member_casual == 'member')
Febuary2023_Casual <- Febuary_2023 %>%    ##Q1 2023
  filter(member_casual == 'casual')
Febuary2023_Member <- Febuary_2023 %>%
  filter(member_casual == 'member')
March2023_Casual <- March_2023 %>%
  filter(member_casual == 'casual')
March2023_Member <- March_2023 %>%
  filter(member_casual == 'member')

December2022_Casual <- December_2022 %>%
  filter(member_casual == 'casual')
December2022_Member <- December_2022 %>%
  filter(member_casual == 'member')
November2022_Casual <- November_2022 %>%       ##Q4 2022
  filter(member_casual == 'casual')
November2022_Member <- November_2022 %>%
  filter(member_casual == 'member')
October2022_Casual <- October_2022 %>%
  filter(member_casual == 'casual')
October2022_Member <- October_2022 %>%
  filter(member_casual == 'member')

July2022_Casual <- July_2022 %>%
  filter(member_casual == 'casual')
July2022_Member <- July_2022 %>%
  filter(member_casual == 'member')
August2022_Casual <- August_2022 %>%
  filter(member_casual == 'casual')
August2022_Member <- August_2022 %>%      ##Q3 2022
  filter(member_casual == 'member')
September2022_Casual <- September_2022 %>%
  filter(member_casual == 'casual')
September2022_Member <- September_2022 %>%
  filter(member_casual == 'member')

```

##### Used RBIND to make tables for Q3 & Q4 of 2022 casual riders and members, and Q1 & Q2 of 2023. I realize 
##### there are faster ways to get here.

```{r}
Q3_2022_Casual <- rbind(July2022_Casual, August2022_Casual, September2022_Casual)
Q3_2022_Member <- rbind(July2022_Member, August2022_Member, September2022_Member)
Q4_2022_Casual <- rbind(October2022_Casual, November2022_Casual, December2022_Casual)
Q4_2022_Member <- rbind(October2022_Member, November2022_Member, December2022_Member)
Q1_2023_Casual <- rbind(January2023_Casual, Febuary2023_Casual, March2023_Casual)
Q1_2023_Member <- rbind(January2023_Member, Febuary2023_Member, March2023_Member)
Q2_2023_Casual <- rbind(April2023_Casual, May2023_Casual, June2023_Casual)
Q2_2023_Member <- rbind(April2023_Member, May2023_Member, June2023_Member)

```

##### Adding a ride length column in each to observe the amount of time both rider types take.Total time /60 so we can see length in minutes.

```{r}

Q2_2023_Casual <- Q2_2023_Casual %>%
  mutate(ride_length = (ended_at - started_at)/60)

```

##### Calculating the mean ride length of different rider types in each quarter.


summarise(Q3_2022_Casual, mean(ride_length))     ## 29.30 minutes avg 
summarise(Q3_2022_Member, mean(ride_length))     ## 13.49 minutes avg
summarise(Q4_2022_Casual, mean(ride_length))     ## 24.42 minutes avg
summarise(Q4_2022_Member, mean(ride_length))     ## 11.43 minutes avg
summarise(Q1_2023_Casual, mean(ride_length))     ## 22.35 minutes avg
summarise(Q1_2023_Member, mean(ride_length))     ## 10.5 minutes avg
summarise(Q2_2023_Casual, mean(ride_length))     ## 28.7 minutes avg
summarise(Q2_2023_Member, mean(ride_length))       



##### Unexpectedly noticing casual riders ride longer than members. 
##### Members kick off the busy season in Q2 with >1,000,000 rides, Q3 recorded >1,000,000 rides for both members and casual riders. 


